Towards Detecting Stalkerware Through Network Traces

### 2019-03-22 paper/main.tex
edit format
Abstract changes

\begin{abstract}
Spyware is known as unwanted software that sends user state and information to 
an external party. Often these apps are not inherently malicious to the 
operating system and are thus difficult to detect. We explore a subset of 
spyware that are available in the Android Play store intended for spying on 
another person though his or her cell phone and create a new method of 
identifying their presence through network signatures.  
\end{abstract}

@@

\section{Introduction}
In a world where technology now pervades every aspect of our lives, some 
aspects of our relationships with intimate partners have changed too. 
Specifically, when these relationships become abusive, there are more options 
available to the abuser for attempting to retain control over their partner. 
One of these options is the ability to install monitoring software on their 
partner's phone or computer to record their online activity, communications, 
and location. We refer to such software as ``stalkerware.''

>>>>

\begin{abstract}
Intimate partner violence (IPV) is a pervasive problem that has been changed by 
the advent of the digital age. The ubiquity of smartphones has made it easy 
even for someone with no technical skills to perform surveillance of a device 
that they have access to. In the context of IPV, abusers often have access to 
the devices of their partners, and either know or can coerce the passwords to 
those devices. With such unrestricted access, an abundance of surveillance 
options exist, 
whether the goal is to track location, remotely block incoming calls or 
messages, record all communications, monitor browsing history, or lock a 
survivor out of their accounts. For some types of information, no extra 
software is required to gather it. For example, Facebook can reveal a user's 
location to their friends. Another example is Apple's backup service, which can 
make copies of text messages and call logs accessible to anyone with the iCloud 
credentials of the phone. Other types of control that might be attractive to an 
abuser, such as the ability to block incoming communication from certain 
people, require software to be installed on the target device. We refer to such 
software as ``stalkerware.'' Stalkerware is dissimilar to other forms of 
malware because of how easily the adversary can access the target device. 
Consequently, traditional assumptions about security do not apply to it, and 
traditional methods of studying it are ineffective. Antivirus, which can 
normally provide information about malware, is poor at detecting 
stalkerware.\cite{chatterjee_spyware_2018} Network Intrusion Detection Systems 
(IDSs) also have low detection rates. This lack of information results in many 
unanswered questions about the use of stalkerware in the context of IPV.

This work takes a step towards answering one of those questions: namely, 
\textit{how prevalent is stalkerware today?} We identify the traffic signatures 
of 24 stalkerware apps, with the goal of being able to identify it from a 
network operator's vantage point. We show that our signatures identify 
stalkerware with few false positives. Additionally, we present a simple method 
of generalizing the signatures we identify, and show that our generalized 
signatures also give few false positives. We also show that peculiarities of 
the stalkerware ecosystem may aid researchers in detecting it, and in 
generalizing specific network signatures to apps that have not been manually 
identified. The ethical and bureaucratic issues inherent in doing a large-scale 
measurement study on real traffic prevented us from using our signatures to 
estimate how much stalkerware exists in the wild. However, we present a plan 
for how we could move forward. 
\end{abstract}

@@

\section{Introduction}
One in four women and one in six men will be subject to some sort of intimate 
partner abuse during their lifetimes, where ``intimate partner abuse'' can 
include physical, emotional, or sexual abuse, as well as stalking or rape. In a 
world where technology now pervades every aspect of our lives, some 
aspects of IPV have recently changed too. Specifically, there are more options 
available to the abuser for attempting to retain control over their partner. 
One of these options is the ability to install monitoring software on their 
partner's phone or computer to record their online activity, communications, 
and location. We refer to such software as ``stalkerware.''
